Global phase diagram and possible quantum spin liquid in the triangular $J-K-\Gamma$ model

### 2019-12-30 paper/paper.tex
20191230 18:53:30

\documentclass[aps,prb,reprint,groupedaddress,showpacs,amsfonts,amsmath,amssymb,superscriptaddress]{revtex4-1}
\usepackage{graphicx}
% \usepackage{subfigure}
\usepackage{bm}
\usepackage{color}
% \usepackage{changes}
    % \definechangesauthor[name={Wang Shi}, color=orange]{swang}

@@

On the other hand, exchange frustation in systems with strongly anisotropic magnetic interactions has been shown to be another promising approach to explore exotic quantum spin states. Like geometric frustation, the effect of exchange frustration is to prevent the formation of long range magnetic order and given raise to a residual ground-state entropy. The spin-$1/2$ Kitaev model \cite{Kitaev2006} on honeycomb lattice, which has both gapped and gapless quantum spin liquid (QSL) ground state (GS) supporting fractionalized excitations, is an example of a model with exchange frustration. As pointed out by Jackeli and Khaliullin \cite{Khaliullin2005, PhysRevLett.102.017205}, the elementary ingredients for realizing this highly anisotropic spin model are strong relativistic spin-orbit coupling (SOC), electron interactions, and specific geometric structure: edge sharing octahedra with 90$^\circ$ TM-O-TM bond angles. Because of its theoretical importance and potential application in quantum computing, great efforts have been made to search a solid-state realization of the Kitaev model. It was first proposed to be realized in iridates $A_2$IrO$_3$ (A = Na or Li) \cite{PhysRevLett.105.027204,PhysRevLett.108.127204,Chun2015,PhysRevLett.110.076402,PhysRevB.87.220407,PhysRevLett.110.097204,PhysRevLett.108.127203,PhysRevB.92.024413,PhysRevLett.112.077204,Rau2014}, and then turned towards $\alpha$-RuCl$_3$ \cite{PhysRevB.91.180401,Banerjee2016,PhysRevB.92.235119,PhysRevB.94.161106,PhysRevB.93.214431,PhysRevLett.118.107203,PhysRevLett.118.107203,PhysRevB.96.115103} in which Ir$^{4+}$/Ru$^{3+}$ ions are arranged in a honeycomb lattice and carry effective $j_{eff}=1/2$ moments. In fact, magnetic ions located at the center of edge-shared octahedra can not only form honeycomb lattice, but also triangular lattice and the Kitaev term can naturally be generalized to this scenario \cite{PhysRevB.93.104417,PhysRevB.89.014414}. Moreover, the symmetric off-diagonal $\Gamma$ term is allowed by symmetry and is a generic feature of S=$1/2$ models with edge sharing octahedra \cite {PhysRevLett.112.077204}. Kai Li \emph{et al.,} \cite{KaiLi2015} have studied the Heisenberg-Kitaev (HK) model on triangular lattice and a global phase diagram with a mean-field level chiral spin liquid (SL) as well as four magnetically ordered phases has been obtained. However, the effect of the $\Gamma$ term on triangluar lattice has not gain as much attention as on honeycomb lattice \cite{PhysRevLett.112.077204,Rau2014,PhysRevLett.118.107203,PhysRevB.96.115103,PhysRevB.93.214431,PhysRevB.100.144422}.

@@

Inspired by previous theoretical and experimental works, we study the Heisenberg-Kitaev-$\Gamma$ ($J-K-\Gamma$) model on the triangular lattice. The $J-K-\Gamma$ model Hamiltonian is given by
\begin{equation}
    H=\sum_{\langle i,j \rangle \in \alpha \beta (\gamma)} \lbrack J \bm{S}_i \bm{\cdot} \bm{S}_j + K S_i^{\gamma} S_j^{\gamma} + \Gamma (S_i^{\alpha} S_j^{\beta} + S_i^{\beta} S_j^{\alpha}) \rbrack
\end{equation}
where $\langle i,j \rangle$ denotes the NN bonds, $\gamma$ takes value $x$, $y$, or $z$ depending on the direction of the NN bond as shown in Fig. 1, and $\alpha$, $\beta$ are the remaining directions. $J$ and $K$ are the magnitude of the Heisenberg and Kitaev interactions and $\Gamma$ the symmetric off-diagonal exchanges. To the best of our knowledge, no exact solution has been reported so far for the spin-$1/2$ Kitaev and/or $\Gamma$ model on the triangular lattice. Therefore, it remains conceptually interesting to investigate whether the QSL could exist as a possible ground state of the triangular $J-K-\Gamma$ model.

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig1.pdf}
    \caption{(Color online) (a) Top view of the triangular lattice of the edge-sharing octahedral. (b) Three different direction of the NN bonds on the triangular lattice, namely $\gamma=x, y, z$ colored red, green and blue respectively. The different colors of the lattice sites label the four sublattices realizing the four-sublattice-transformation. (c) The orientation of the cubic $x$, $y$, $z$ with respect to the octahedral. The spin operators $S^x$, $S^y$ and $S^z$ are defined with respect to this reference frame.}
     \label{fig:ModelDefinition}
\end{figure}

\section{Methods}

The ground state of the full anisotropic Hamiltonian is no longer solvable in closed analytical form, even in the classical case. One, therefore, either has to rely on approximate analytical method or on numerical techniques. Here, we briefly describe the methods we used. \textcolor{red}{$\cdots\cdots$}

\subsection{Dual Transformation}
\textcolor{red}{A introduction to dual transformation. Left blank currently.}

@@

\subsection{Exact Diagonalization}
\textcolor{red}{A introduction to exact diagonalization. Left blank currently.}

\section{Phase Diagram and Analysis}
\textcolor{red}{Global phase diagram of the $J-K-\Gamma$ model. Left blank currently.}

\begin{figure*}
    \centering
    % \subfigure{\includegraphics[width=0.96\textwidth]{Fig2_a.pdf}}\\
    % \subfigure{\includegraphics[width=0.96\textwidth]{Fig2_b_f.pdf}}\\
    \includegraphics[width=0.96\textwidth]{Fig2.pdf}
    \caption{(Color online) (a) Global phase diagram of the triangular lattice J-K-$\Gamma$ model. The angle $\alpha$ and $\beta$ denote the radial and azimuthal angles, respectively. There are in total ten phases including three different FM phases denoted as FM-A, FM-B and FM-C, four different stripe phases designated as Stripe-A, Stripe-B, Stripe-C and Stripe-D, 120$^\circ$ N\'{e}el, Dual N\'{e}el and a possible quantum spin liquid phase. (b) Illustration of the FM order. All spins along the same direction. (c) Illustation of the stripe order of the $x$ type. That is the spins along the x-bond direction are parallel and adjacent chains are antiparallel. There are also stripe order of the $y$ or $z$ type, which are degenerate with the $x$ type stripe order. (d) Illustration of the 120$^\circ$ N\'{e}el order. All spins are coplanar and the angles between nearest-neighbor spins are 120$^\circ$. (e) Illustration of the Dual N\'{e}el order. The magnetic unit cell include 12 lattice sites as marked by the green dashed parallelogram.}
     \label{fig:PhaseDiagram}
\end{figure*}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig4.pdf}
    \caption{(Color online) Map of the probabilities.}
     \label{fig:StructureFactors}
\end{figure}

>>>>

\documentclass[aps,prb,reprint,groupedaddress,showpacs,amsfonts,amsmath,amssymb,superscriptaddress]{revtex4-1}
\usepackage{IEEEtrantools}
\usepackage{graphicx}
% \usepackage{subfigure}
\usepackage{bm}
\usepackage{color}
% \usepackage{changes}
    % \definechangesauthor[name={Wang Shi}, color=orange]{swang}

@@

On the other hand, exchange frustation in systems with strongly anisotropic magnetic interactions has been shown to be another promising approach to explore exotic quantum spin states. Like geometric frustation, the effect of exchange frustration is to prevent the formation of long range magnetic order and given raise to a residual ground-state entropy. The spin-$1/2$ Kitaev model \cite{Kitaev2006} on honeycomb lattice, which has both gapped and gapless quantum spin liquid (QSL) ground state (GS) supporting fractionalized excitations, is an example of a model with exchange frustration. As pointed out by Jackeli and Khaliullin \cite{Khaliullin2005, PhysRevLett.102.017205}, the elementary ingredients for realizing this highly anisotropic spin model are strong relativistic spin-orbit coupling (SOC), electron interactions, and specific geometric structure: edge sharing octahedra with 90$^\circ$ TM-O-TM bond angles. Because of its theoretical importance and potential application in quantum computing, great efforts have been made to search a solid-state realization of the Kitaev model. It was first proposed to be realized in iridates $A_2$IrO$_3$ (A = Na or Li) \cite{PhysRevLett.105.027204,PhysRevLett.108.127204,Chun2015,PhysRevLett.110.076402,PhysRevB.87.220407,PhysRevLett.110.097204,PhysRevLett.108.127203,PhysRevB.92.024413,PhysRevLett.112.077204,Rau2014}, and then turned towards $\alpha$-RuCl$_3$ \cite{PhysRevB.91.180401,Banerjee2016,PhysRevB.92.235119,PhysRevB.94.161106,PhysRevB.93.214431,PhysRevLett.118.107203,PhysRevLett.118.107203,PhysRevB.96.115103} in which Ir$^{4+}$/Ru$^{3+}$ ions are arranged in a honeycomb lattice and carry effective $j_{eff}=1/2$ moments. In fact, magnetic ions located at the center of edge-shared octahedra can not only form honeycomb lattice, but also triangular lattice (see Fig. \ref{fig:ModelDefinition}(a))and the Kitaev term can naturally be generalized to this scenario \cite{PhysRevB.93.104417,PhysRevB.89.014414}. Moreover, the symmetric off-diagonal $\Gamma$ term is allowed by symmetry and is a generic feature of S=$1/2$ models with edge sharing octahedra \cite {PhysRevLett.112.077204}. Kai Li \emph{et al.,} \cite{KaiLi2015} have studied the Heisenberg-Kitaev (HK) model on triangular lattice and a global phase diagram with a mean-field level chiral spin liquid (SL) as well as four magnetically ordered phases has been obtained. However, the effect of the $\Gamma$ term on triangluar lattice has not gain as much attention as on honeycomb lattice \cite{PhysRevLett.112.077204,Rau2014,PhysRevLett.118.107203,PhysRevB.96.115103,PhysRevB.93.214431,PhysRevB.100.144422}.

@@

Inspired by previous theoretical and experimental works, we study the Heisenberg-Kitaev-$\Gamma$ ($J-K-\Gamma$) model on the triangular lattice. The $J-K-\Gamma$ model Hamiltonian is given by
\begin{equation}
    H=\sum_{\langle i,j \rangle \in \alpha \beta (\gamma)} \lbrack J \bm{S}_i \bm{\cdot} \bm{S}_j + K S_i^{\gamma} S_j^{\gamma} + \Gamma (S_i^{\alpha} S_j^{\beta} + S_i^{\beta} S_j^{\alpha}) \rbrack
    \label{eq:Hamiltonian}
\end{equation}
where $\langle i,j \rangle$ denotes the NN bonds, $\gamma$ takes value $x$, $y$, or $z$ depending on the direction of the NN bond as shown in Fig.\ref{fig:ModelDefinition}(b), and $\alpha$, $\beta$ are the remaining directions. $J$ and $K$ are the magnitude of the Heisenberg and Kitaev interactions and $\Gamma$ the symmetric off-diagonal exchanges. To the best of our knowledge, no exact solution has been reported so far for the spin-$1/2$ Kitaev and/or $\Gamma$ model on the triangular lattice. Therefore, it remains conceptually interesting to investigate whether the QSL could exist as a possible ground state of the triangular $J-K-\Gamma$ model.

\begin{figure}
    \includegraphics[width=\columnwidth]{Fig1.pdf}
    \caption{(Color online) (a) Top view of the triangular lattice of the edge-sharing octahedron. (b) Three types of the NN bonds on the triangular lattice, namely $\gamma=x, y, z$ colored red, green and blue, respectively. The different colors of the lattice sites label the four sublattices realizing the four-sublattice-transformation. (c) The orientation of the cubic $x$, $y$, $z$ axes with respect to the octahedron. The spin operators $S^x$, $S^y$ and $S^z$ are defined with respect to this reference frame.}
    \label{fig:ModelDefinition}
\end{figure}

@@

\section{Methods}
The full anisotropic Hamiltonian \eqref{eq:Hamiltonian} is no longer exactly solvable, even for the classical case. One, therefore, either has to rely on approximate analytical method or on numerical techniques. To determine the ground state of the Hamiltonian \eqref{eq:Hamiltonian}, we employ a combination of classical Monte Cralo simulation and ED calculation. Here, we briefly describe the methods we used. For convenience, we fix the energy scale with $\sqrt{J^2 + K^2 + \Gamma^2}=1$ and parametrize the exchanges using angles $\alpha$ and $\beta$
\begin{equation}
    J=\sin\alpha \sin\beta, \quad K=\sin\alpha \cos\beta, \quad \Gamma=\cos\alpha
    \label{eq:Parameters}
\end{equation}
where $\alpha \in [0, \pi]$ and $\beta \in [0, 2\pi]$ to cover the global phase diagram.

@@

\subsection{Exact Diagonalization}
We perform a Lanczos ED calculation of the GS energy of the Hamiltonian \eqref{eq:Hamiltonian} on a 24-site cluster with periodic boundary conditions. To detect quantum phase transitions, the second derivatives of the ground-state energy, $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ were computed and looked for singular features that indicate changes in the groun-state characteristics. Phases containing exactly solvable or well-understood points, such as the FM, 120$^\circ$ N\'{e}el, stripe and Dual N\'{e}el can be readily identified. The remaining phases were identified mainly by examining the static structure factor
\begin{equation}
    S(\mathbf{Q}) = \frac{1}{N} \sum_{ij}e^{i\mathbf{Q}\cdot(\mathbf{r}_i - \mathbf{r}_j)} \langle \mathbf{S}_i \cdot \mathbf{S}_j \rangle
    \label{eq:StaticStructureFactor}
\end{equation}

To extract the moment direction of a magnetic ordered phase from a cluster ground state, we employ the method developed by Ji\v{r}\'{i} Chaloupka \emph{et  al} \cite{PhysRevB.94.064435}. The basic idea of this method is to measure the probabilities of the exact cluster ground state $|GS\rangle$ on cluster spin-coherent states with varying moment directions. The cluster spin-coherent state is a direct product of spin-1/2 coherent states on each site $i$
\begin{equation}
    |\Psi\rangle = \prod_{i=1}^N|\theta_i,\phi_i\rangle
    \label{eq:ClusterCoherentState}
\end{equation}
where the spin-1/2 coherent state
\begin{equation}
    |\theta, \phi \rangle = \mathcal{R}_z(\phi) \mathcal{R}_y(\theta) |\uparrow \rangle = e^{-i\phi S^z} e^{-i\theta S^y} |\uparrow\rangle
    \label{eq:Spin-1/2CoherentState}
\end{equation}
is fully polaried along the $(\theta, \phi)$ direction. Here the cubic axes are used (see Fig. \ref{fig:ModelDefinition}(c)), $\theta$ and $\phi$ are the conventional spherical angles, and $S^z |\uparrow \rangle = \frac{1}{2}|\uparrow \rangle$. By calculating the overlap with the exact cluster GS from ED, and maximizing the probability $P = |\langle \Psi | GS \rangle|^2$ with respect to $\theta$s and $\phi$s, we can then identify the classical pattern that best fits the exact GS.

\section{Classical Analysis}
To understand the effects of inculding this off-diagonal $\Gamma$ term, we first study the classical $J-K-\Gamma$ model where the spin operators are viewed as unit-vectors in three dimension. On our proceeding, we note that the global phase diagram has been obtained previously by using a combination of Luttinger-Tsiza (LT) and classical Monte Carlo simulations \cite{PhysRevB.92.165108}. Six different magnetic orderings: ferromagnet (FM), stripe, 120$^\circ$ N\'{e}el order, nematic, $\mathbb{Z}_2$ and dual-$\mathbb{Z}_2$ vortex crystal phases were found, see Fig. 2 in Ref \onlinecite{PhysRevB.92.165108}. Our classical Monte Carlo study give qualitatively the same phase diagram.

\subsection{Moment direction of the classcial FM order}
For the classical FM order, the energy of the $J-K-\Gamma$ model per lattice site is given by
\begin{equation}
    E_{FM}^{c} = (3J + K) + 2\Gamma(S^xS^y + S^yS^z + S^zS^x)
    \label{eq:EcFM}
\end{equation}
On this level, the moment direction of the classical FM state is determined solely by $\Gamma$ and the problem becomes finding the minimum and maximum of the multi-variable function $f(S^x, S^y, S^z) = S^xS^y + S^yS^z + S^zS^x$ with the constraint $|\mathbf{S}| = 1$. $f(S^x, S^y, S^z)$ takes maximum value $f_{max}=1$ at $S^x=S^y=S^z=\pm 1/\sqrt{3}$ and minimum value $f_{min}=-0.5$ when the conditions $S^x + S^y + S^z = 0$ and $|\mathbf{S}| = 1$ are fulfilled. That is to say, for $\Gamma > 0$ the moment of the classcial FM state prefers to lie in the plane that is perpendicular to the $[111]$ direction and when $\Gamma$ is negative, the moment would along the $[111]$ or $[\bar{1}\bar{1}\bar{1}]$ direction.

\subsection{Moment direction of the classcial stripe order}

In the case of stripe order, there are three degenerate spin configurations as illustrated in Fig. \ref{fig:PhaseDiagram}(b)-(d). For the spin configuration shown in Fig. \ref{fig:PhaseDiagram}(b), the energy of the $J-K-\Gamma$ model Hamiltonian is
\begin{IEEEeqnarray}{rCl}
    E_{StripeX}^{c} & = & -(J + K) + 2KS^xS^x \nonumber \\
        && + 2\Gamma(S^yS^z - S^xS^y - S^zS^x)
    \label{eq:EcStripeX}
\end{IEEEeqnarray}
and the energy for spin configuration in Fig. \ref{fig:PhaseDiagram}(c) and \ref{fig:PhaseDiagram}(d) can be obtained by a cyclic permutation $x \rightarrow y \rightarrow z \rightarrow x$. The moment direction of the classical stripe order is determined by the anisotropy parameters $K$ and $\Gamma$. In general, $E_{StripeX}^{c}$ has three extreme points where the first derivatives equal zero
\begin{align}
    &\mathbf{S}_0: \quad S_{0}^{y}=-S_{0}^{z} = \pm1/\sqrt{2}, \quad S_{0}^{x} = 0\\
    &\mathbf{S}_1: \quad S_{1}^{y}=S_{1}^{z} = f_{1}(K, \Gamma), \quad S_{1}^{x} = g_{1}(K, \Gamma) S_{1}^{y}\\
    &\mathbf{S}_2: \quad S_{2}^{y}=S_{2}^{z} = f_{2}(K, \Gamma), \quad S_{2}^{x} = g_{2}(K, \Gamma) S_{2}^{y}
\end{align}
see Appendix A for their explicit expressions.

For $\Gamma=0, K>0$, the stripe order shown in Fig. \ref{fig:PhaseDiagram}(b) prefers to lie in the $yz$ plane. A definite positive $\Gamma$ would fix the ordered moment to the $\mathbf{S}_0$ direction and negative $\Gamma$ drives it to the $\mathbf{S}_1$ direction. When $\Gamma=0, K<0$, the moment would along the $x$ axis to have lowest energy and both positive and negative $\Gamma$ make it deviate from the $x$ axis and point the $\mathbf{S}_1$ direction.

\textcolor{red}{$\cdots \cdots$}

\section{Global Phase Diagram}
\begin{figure*}
    \centering
    \includegraphics[width=0.96\textwidth]{Fig2.pdf}
    \caption{(Color online) (a) Global phase diagram of the triangular lattice J-K-$\Gamma$ model. The angle $\alpha$ and $\beta$ denote the radial and azimuthal angles, respectively. There are in total ten phases including three different FM phases denoted as FM-A, FM-B and FM-C, four different stripe phases designated as Stripe-A, Stripe-B, Stripe-C and Stripe-D, 120$^\circ$ N\'{e}el, Dual N\'{e}el and a possible quantum spin liquid phase. (b) Illustration of the FM order. All spins along the same direction. (c) Illustation of the stripe order of the $x$ type. That is the spins along the x-bond direction are parallel and adjacent chains are antiparallel. There are also stripe order of the $y$ or $z$ type, which are degenerate with the $x$ type stripe order. (d) Illustration of the 120$^\circ$ N\'{e}el order. All spins are coplanar and the angles between nearest-neighbor spins are 120$^\circ$. (e) Illustration of the Dual N\'{e}el order. The magnetic unit cell include 12 lattice sites as marked by the green dashed parallelogram.}
     \label{fig:PhaseDiagram}
\end{figure*}

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{Fig4.pdf}
    \caption{(Color online) Map of the probabilities.}
     \label{fig:Proabilities}
\end{figure}

\textcolor{red}{Global phase diagram of the $J-K-\Gamma$ model. Left blank currently.}
